Transferability of clinical prediction models for early trauma care

### 2019-07-01 ManuscriptMarkdown.Rmd

## LOAD FUNCTIONS, ENVIROMENT AND TABLES
  ## Source functions
  source("functions/WorkTableCreator.R")
  ## Original results
  results <- readRDS("output/original.results.Rds")
  ## Compiled results with UI as data frame
  result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
  ## Create table for ease of data retrival
  results.all.data <- WorkTableCreator(result.tables.data.frame)

@@

## TABLE SETUP
  # Table 1 = Sample characteristics table
  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
  names(high.low) <- c("characteristic", "level", "high volume", "low volume")
  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
  names(met.nonmet) <- c("characteristic", "level", "metropolitan", "non metropolitan")
  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
  names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
  # Table 2 = result.tables, cropped to only show validation
  table.2 <- results.all.data[1:3,]
  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 3 = Cropped to only show transfer
  table.3 <- results.all.data[4:6,]
  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 4 = Cropped to only show transfer differance
  table.4 <- results.all.data[7:9,]
  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
```
---
abstract: |

@@

  **Methods:** Patients age 15 or above registered between the years 2011 and
  2016 in the Swedish national trauma registry SweTrau were included. Three sets
  of data were created: High and low volume centres, metropolitan and
  non-metropolitan centres, and multi and single centre data. Clinical
  prediction models based on logistic regression were developed in each set of
  data and transferred within sets. Model performance was evaluated in all data
  sets using mistriage rate, undertriage rate and overtriage rate. Multiple
  imputation using chained equation was used to handle missing data. Model
  performance was reported as medians with 95% uncertainty interval across
  imputed data sets.

  **Results:** `r total.number.of.patients ` patients were included. Model
  validation median mistriage rate with 95% UI ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
  across all samples. Transfer performance when compared to receiving sample validation 
  performance resulted in subtantial uncertainty in all samples. Worst comparison 
  performance after transfer was found in the `r colnames(highest.comp.mistriage)` 
  sample with a median mistriage difference of `r highest.comp.mistriage`.

  **Conclusions:** In all samples, model transfer was associated with high 
  levels of uncertainty. Notably, model transfer from the `r colnames(highest.comp.mistriage)` 
  sample to the other sample in the data set lead to an increased mistriage
  rate, mainly due to a marked increase in overtriage.

@@

![](fig1.jpg)  
**Figure 1: High- and low volume centre data set.** Initial split based on
number of patients. Temporal split made using date of traumatic event.

@@

**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
```{r, echo = FALSE, results = "asis"} 
## Table 1
options(knitr.kable.NA = '')
kable(high.low, format = "markdown")
kable(met.nonmet, format = "markdown")
kable(multi.single, format = "markdown")
``` 
Abbreviations: GCS glasgow coma scale, SBP systolic blood pressure, RR respiratory rate, ISS injury severity score, NISS new injury severity score

@@

**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).
```{r, echo = FALSE, results = "asis"} 
## Table 2
kable(table.2[,1:2], format = "markdown")
kable(table.2[,3:4], format = "markdown")
kable(table.2[,5:6], format = "markdown")
```

>>>>

## LOAD FUNCTIONS, ENVIROMENT AND TABLES
  ## Source functions
  source("functions/WorkTableCreator.R")
  ## Retrive original results
  results <- readRDS("output/original.results.Rds")
  ## Retrive data frame with compiled results
  result.tables.data.frame <- readRDS("result.tables.data.frame.Rds")
  ## Create table from result.tables.data.frame for ease of data retrival in document
  results.all.data <- WorkTableCreator(result.tables.data.frame)

@@

## TABLE SETUP
  # Table 1 = Sample characteristics table
  high.low <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
  names(high.low) <- c("characteristic", "level", "high volume", "low volume")
  met.nonmet <- results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan
  names(met.nonmet) <- c("characteristic", "level", "metropolitan", "non metropolitan")
  multi.single <- results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre
  names(multi.single) <- c("characteristic", "level", "multi centre", "single centre")
   # Create complete table one (test)
   complete.table.one <- results$table.one.list$Characteristics_table_of_high.volume.vs.low.volume
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_metropolitan.vs.non.metropolitan$non.metropolitan)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$multi.centre)
   complete.table.one <- cbind(complete.table.one, results$table.one.list$Characteristics_table_of_multi.centre.vs.single.centre$single.centre.11001)
   names(complete.table.one) <- c("characteristic", "level", "high volume", "low volume", "metropolitan", "non metropolitan", "multi centre", "single centre")
  # Table 2 = result.tables, cropped to only show validation
  table.2 <- results.all.data[1:3,]
  rownames(table.2) <- c("Mistriage", "Undertriage", "Overtriage")
   # Create complete table two (test) 
   complete.table.two <- results.all.data[1:3,]
   rownames(complete.table.two) <- c("mistriage", "undertriage", "overtriage")
  # Table 3 = Cropped to only show transfer
  table.3 <- results.all.data[4:6,]
  rownames(table.3) <- c("Mistriage", "Undertriage", "Overtriage")
  # Table 4 = Cropped to only show transfer differance
  table.4 <- results.all.data[7:9,]
  rownames(table.4) <- c("Mistriage", "Undertriage", "Overtriage")
```
---
abstract: |

@@

  **Methods:** Patients age 15 or above registered between the years 2011 and
  2016 in the Swedish national trauma registry SweTrau were included. Three sets
  of data were created: High and low volume centres, metropolitan and
  non-metropolitan centres, and multi and single centres. Clinical
  prediction models were developed usinglogistic regression in each data set 
  and transferred within sets. Model performance was evaluated in all data
  sets using mistriage rate, undertriage rate and overtriage rate. Multiple
  imputation using chained equation was used to handle missing data. Model
  performance was reported as medians with 95% uncertainty interval across
  imputed data sets.

  **Results:** `r total.number.of.patients ` patients were included. Model
  validation mistriage rate ranged from `r lowest.val.mistriage` 
  to `r highest.val.mistriage`, and transferred model
  mistriage rate ranged from `r lowest.tran.mistriage` to `r highest.tran.mistriage` 
  across all samples. Performance after transfer when compared to receiving sample validation 
  performance resulted in subtantial uncertainty in all samples. Worst performance after 
  transfer was found in the `r colnames(highest.comp.mistriage)` 
  sample with a mistriage difference of `r highest.comp.mistriage`.

  **Conclusions:** In all samples model transfer was associated with high 
  levels of uncertainty. Notably, transfer from the `r colnames(highest.comp.mistriage)` 
  sample to the other sample in the data set lead to an increased mistriage
  rate, primarily due to increased overtriage.

@@

![](fig1.jpg)  
**Fig. 1: High- and low volume centre data set.** Initial split based on
number of patients. Temporal split made using date of traumatic event.

@@

##### page break

**Table 1. Sample characteristics of multiple imputed data.** Data is presented as medians with interquartile range (IQR) or counts with % as applicable.
```{r, echo = FALSE, results = "asis"} 
## Table 1
options(knitr.kable.NA = '')
# kable(high.low, format = "markdown")
# kable(met.nonmet, format = "markdown")
# kable(multi.single, format = "markdown")
kable(complete.table.one, format = "markdown")
``` 
GCS, glasgow coma scale; SBP, systolic blood pressure; RR, respiratory rate; 30DS, 30 day survival; ISS, injury severity score; NISS, new injury severity score

@@

**Table 2. Model validation performance, i.e. performance when applied to the sample in which the model was created.** Data is presented as point estimates with 95% Uncertainty Intervals (95% UI).
```{r, echo = FALSE, results = "asis"} 
## Table 2
# kable(table.2[,1:2], format = "markdown")
# kable(table.2[,3:4], format = "markdown")
# kable(table.2[,5:6], format = "markdown")
kable(complete.table.two, format = "markdown")
```
